---
title: "Retirement Variables"
author: "Giovanni"
date: "2025-05-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(haven)
library(dplyr)
library(tidyr)
library(stringr)

current_wave <- 16  # change this to the number of waves
                    # in the dataset you are using 
                    # 16 is for the year 2022

df <- read_dta("randhrs1992_2022v1.dta")
```


**RwIEARN** : Sum or respondent's wages including all sources of income
              w stands for wave


## Selecting Variables of Interest

```{r}
constants <- c("hhidpn")  # Put variables constant over time in here
rcolumns <- c() 
var_cols <- c("r%diearn") # Put variables that change each wave 
                          # with `%d` instead of wave as a string

# note if using the datasets published by the HRS variables 
# are stored in all caps while RAND using lowercase


for (number in 1:current_wave) {
  for (col in var_cols) {
    rcolumns <- c(rcolumns, sprintf(col, number))
  }
}


df_sub <- df %>%
  select(all_of(constants), all_of(rcolumns)) # subset of original data 
                                              # with variables of interest
                                              # maintaining HRS convention

```





```{r}

name_df <- data.frame(Question = character(),
                 Wave = integer()
                 )

# create a df 'key' that we can join later
for (number in 1:current_wave) {
  for (col in var_cols) {
    name_df <- rbind(name_df, data.frame(Question = sprintf(col, number), Wave = number))
  }
} 

# create a Wave column each observation can now be defined by 
# `Wave` and `hhidpn`
long_df <- df_sub %>%
  pivot_longer(cols = name_df$Question,
               names_to = "Question",
               values_to = "Response") %>%
  left_join(name_df, by = "Question") %>%
  mutate(Question = str_remove_all(Question, "[0-9]")) %>%
  pivot_wider(names_from = Question,
              values_from = Response)

rm(name_df)
```


The dataset comes from the Health and Retirement Study. Specifically I used the longitudinal file:

Health and Retirement Study, (RAND HRS Longitudinal File 2022 (V1)) public use dataset. Produced
and distributed by the University of Michigan with funding from the National Institute on Aging (grant numbers
NIA U01AG009740 and NIA R01AG073289). Ann Arbor, MI, (May 2025).


RAND HRS Longitudinal File 2022 (V1). Produced by the RAND Center for the Study of Aging, with
funding from the National Institute on Aging and the Social Security Administration. Santa Monica, CA (May
2025).
